---
title: "p8105_hw6_fwt2107"
author: "Felix Tran"
date: "November 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "right"))
```

# Problem 1

### Data cleaning

1. Read in the dataset from the Washington Post's github page

2. Create a **city_state** variable to combine each homicide's city and state

3. Remove all homicides which were recorded in Dallas, TX; Phoenix, AZ; Kansas
City, MO; and Tulsa, AL

4. Transform **victim_age** into a numeric variable

5. Transform **victim_race** into a binary variable (white vs. non-white)

6. Create a **solved_binary** binary variable for the outcome of each homicide
(solved vs. unsolved)
```{r}
data_url <- RCurl::getURL('https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv')

homicide_df <- readr::read_csv(data_url) %>% 
  mutate(city_state = stringr::str_c(city, ', ',state)) %>% 
  filter(!(city_state %in% c('Dallas, TX', 'Phoenix, AZ', 'Kansas City, MO', 
                             'Tulsa, AL'))) %>% 
  mutate(victim_age = as.numeric(victim_age),
         victim_race = ifelse(victim_race == "White", "white", "non-white"),
         solved_binary = ifelse(disposition == "Closed by arrest", "solved", 
                                "unsolved"))
```

```{r}

```

